---
title: "Gelsolin ANOVA with confounders"
author: "Katie Overmyer"
date: "6/5/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message=F, error=F, warning=FALSE}
library(DBI)
library(RSQLite)

#### Establish a connection to the DB
con <- dbConnect(RSQLite::SQLite(), dbname = "P:/All_20200428_COVID_plasma_multiomics/SQLite Database/Covid-19 Study DB.sqlite")

# check
pvalues <- dbGetQuery(con, "SELECT * FROM pvalues WHERE biomolecule_id = 7974")


df_proteins<- dbGetQuery(con, "SELECT deidentified_patient_metadata.sample_id, normalized_abundance, proteomics_measurements.biomolecule_id, COVID, Hospital_free_days_45, Age_less_than_90, Gender, ICU_1, DM, Charlson_score, SOFA
           FROM proteomics_measurements
           INNER JOIN proteomics_runs ON proteomics_runs.replicate_id = proteomics_measurements.replicate_id
           INNER JOIN rawfiles ON rawfiles.rawfile_id = proteomics_runs.rawfile_id
           INNER JOIN deidentified_patient_metadata ON deidentified_patient_metadata.sample_id = rawfiles.sample_id
           INNER JOIN biomolecules on biomolecules.biomolecule_id = proteomics_measurements.biomolecule_id
           WHERE rawfiles.keep = 1  
           AND biomolecules.biomolecule_id = '7974'
           ")

# disconnect
dbDisconnect(con) 


df_proteins$Gender[df_proteins$Gender == ""] <- NA

```

## Analysis uing linear regression to test the effect of gelsolin normalized abundance levels on response variable Hospital Free Days at 45. 

### These data suggest Gelsolin is predictive of hospital free days. 

```{r}
fit <- lm(Hospital_free_days_45 ~ normalized_abundance * COVID, data = df_proteins)
anova(fit)

```

### Let's also ask if COVID status has effect on hospital free days:

```{r}

fit2 <- lm(Hospital_free_days_45 ~ COVID, data =df_proteins)
anova(fit2)

```

### The pvalue of 0.003 suggest that COVID status is significant also.

### However, likelyhood ratio test between models suggests that Gelsolin values improves prediction of hospital free days

```{r}
anova(fit, fit2)
```

### If we include confounders of gender and age, Gelsolin levels still have significant effect on hospital free days. 

```{r}

fit_3 <- lm(Hospital_free_days_45 ~ normalized_abundance + COVID + Gender + Age_less_than_90, data = df_proteins)

anova(fit_3)

```

